---
title: "Rare proj strings to WKT"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Background

[CRS](https://www.nceas.ucsb.edu/sites/default/files/2020-04/OverviewCoordinateReferenceSystems.pdf) (aka coordinate reference system) is how spatial data knows what projection it is in (if any).  For a long time [rgdal](https://CRAN.R-project.org/package=rgdal) has been the intermediary between the many CRSs in the world and R.  It has primarily supported [Proj strings](https://proj.org/) to negotiate data to-from R and [GDAL](https://gdal.org/).

Changes are afoot.  Soon (2023) rgdal maintenance will end in favor of direct interaction between R and GDAL. Soon the necessary coding/decoding will be handled within the [sf](https://CRAN.R-project.org/package=sf) package using [EPSG-like](https://epsg.io/) codes or [Well Known Text](https://en.wikipedia.org/wiki/Well-known_text_representation_of_coordinate_reference_systems#ESRI_vs_OGC) aka WKT.  


### Transitions when it is easy

Many datasets come with simple log/lat values - that is there is no projection.  In this case the easiest CRS is `EPSG:4326` or `WGS84`.  The following are all equivalent.

```{r setup_easy}
library(sf)
crs_a <- sf::st_crs("EPSG:4326")
crs_b <- sf::st_crs(4326)
crs_c <- sf::st_crs("WGS84")
(crs_a == crs_b) && (crs_a == crs_c)
```

### Transitions when it is hard

In the meantime, there is a lot of data served that is projected and that retains to use of Proj strings rather than an EPSG code or WKT.  Below is an example from a NetCDF file of a ocean dynamics model. While we still have access to rdgal package we can leverage it to help convert frpm proj strings to WKT or possibly EPSG (for those stored that might be in your database of CRSs.)

```{r setup_hard}
library(rgdal)
proj_string <- "proj=tmerc +datum=NAD83 +lon_0=-70d10 lat_0=42d50 k=.9999666666666667 x_0=900000 y_0=0"
```

Here we use rgdal to cast the proj_string to WKT

```{r showWKT}
wkt <- rgdal::showWKT(proj_string)
wkt
```

Now we can pass that to `sf::st_crs()` to use within sf package.

```{r st_crs}
crs <- sf::st_crs(wkt)
crs
```

### What happens when rgdal is no longer supported?

Not sure.  Possibly [gdalUtilities](https://github.com/JoshOBrien/gdalUtilities) or [gdalUtils](https://github.com/JoshOBrien/gdalUtils)?
