---
title: "FVCOM"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Convenient access to [FVCOM](http://fvcom.smast.umassd.edu/fvcom/)
datasets from R.

### Background

[FVCOM](http://fvcom.smast.umassd.edu/fvcom/) produces ocean circulation
models on an irregular mesh. Here is the [user
manual](http://fvcom.smast.umassd.edu/wp-content/uploads/2013/11/MITSG_12-25.pdf).

As an example, [NECOFS](http://fvcom.smast.umassd.edu/necofs/) leverages
this mesh to deliver models for the Gulf of Maine, Mass coastal waters
and a small number of estuaries. NECOFS provides an [interactive web
map](http://134.88.228.119:8080/fvcomwms/)

## Requirements

 + [R 4.1+](https://www.r-project.org/)
  + [rlang](https://cran.r-project.org/package=rlang)
  + [ncdf4](https://cran.r-project.org/package=ncfd4)
  + [tibble](https://cran.r-project.org/package=tibble)
  + [dplyr](https://cran.r-project.org/package=dplyr)
  + [raster](https://cran.r-project.org/package=raster)
  + [sf](https://cran.r-project.org/package=sf)
  + [fasterize](https://cran.r-project.org/package=fasterize)
  + [locate](https://cran.r-project.org/package=ncfd4) # Not available on CRAN


#### Gulf of Maine centric products

There are a number of models with a Gulf of Maine focus including
[forecasts](http://www.smast.umassd.edu:8080/thredds/forecasts.html) and
[archives](http://www.smast.umassd.edu:8080/thredds/catalog/models/fvcom/NECOFS/Archive/NECOFS_GOM/catalog.html).  For each of these OPeNDAP resources we have provided a simple interface class.

  + [GOM](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_GOM4_FORECAST.nc.html) `GOMPhysics()`
  
  + [Boston](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_FVCOM_OCEAN_BOSTON_FORECAST.nc.html) `BostonPhysics()`
  
  + [Scituate](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_FVCOM_OCEAN_SCITUATE_FORECAST.nc.html) `ScituatePhysics()`
  
  + [Mass Bay](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_FVCOM_OCEAN_MASSBAY_FORECAST.nc.html) `MassBayPhysics()`
  
  + [Saco Bay](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_FVCOM_OCEAN_SACO_FORECAST.nc.html) `SacoBayPhysics()`
  
  + [Hampton](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_FVCOM_OCEAN_HAMPTON_FORECAST.nc.html) `HamptonPhysics()`
  
  + [Global](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_FVCOM_OCEAN_GLOBAL_FORECAST.nc.html) `GlobalPhysics()`
  



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
