---
title: "FVCOM"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Convenient access to [FVCOM](https://www.fvcom.org/) datasets from R.

# Background

[FVCOM](https://www.fvcom.org/) produces ocean circulation
models on an irregular mesh. Here is the [user
manual](http://fvcom.smast.umassd.edu/wp-content/uploads/2013/11/MITSG_12-25.pdf).

As an example, [NECOFS](http://134.88.228.119:8080/fvcomwms/) leverages
this mesh to deliver models for the Gulf of Maine, Mass coastal waters
and a small number of estuaries. NECOFS provides an THREDDS catalog of its OPeNDAP resources.

# Requirements

 + [R 4.1+](https://www.r-project.org/)
 + [rlang](https://cran.r-project.org/package=rlang)
 + [ncdf4](https://cran.r-project.org/package=ncfd4)
 + [tibble](https://cran.r-project.org/package=tibble)
 + [dplyr](https://cran.r-project.org/package=dplyr)
 + [sf](https://cran.r-project.org/package=sf)
 + [stars](https://cran.r-project.org/package=stars)
 + [locate](https://github.com/BigelowLab/locate) # Not available on CRAN

# Installation

```
remotes::install_github("BigelowLab/locate")`
remotes::install_github("BigelowLab/fvcom")
```


# Gulf of Maine centric products

There are a number of models with a Gulf of Maine focus including
[forecasts](http://www.smast.umassd.edu:8080/thredds/catalog/models/fvcom/NECOFS/Forecasts/catalog.html) and
[hindcasts](http://www.smast.umassd.edu:8080/thredds/catalog/models/fvcom/NECOFS/Archive/Seaplan_33_Hindcast_v1/catalog.html).  For each of these OPeNDAP resources we have provided a simple interface class.

  + [GOM](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_GOM3_FORECAST.nc.html) `GOMPhysics()`
  
  + [Boston](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_FVCOM_OCEAN_BOSTON_FORECAST.nc.html) `BostonPhysics()`
  
  + [Scituate](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_FVCOM_OCEAN_SCITUATE_FORECAST.nc.html) `ScituatePhysics()`
  
  + [Mass Bay](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_FVCOM_OCEAN_MASSBAY_FORECAST.nc.html) `MassBayPhysics()`
  
  + [Saco Bay](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_FVCOM_OCEAN_SACO_FORECAST.nc.html) `SacoBayPhysics()`
  
  + [Hampton](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_FVCOM_OCEAN_HAMPTON_FORECAST.nc.html) `HamptonPhysics()`
  
  + [Global](http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Forecasts/NECOFS_FVCOM_OCEAN_GLOBAL_FORECAST.nc.html) `GlobalPhysics()`
  
# Accessing FVCOM data

```{r}
suppressPackageStartupMessages({
  library(sf) 
  library(fvcom) 
  library(ncdf4) 
  library(dplyr)
})
```

Open the FVCOM resource as you would any NetCDF file.

```{r}
uri_base <- "http://www.smast.umassd.edu:8080/thredds/dodsC/models/fvcom/NECOFS/Archive/Seaplan_33_Hindcast_v1"
uri <- file.path(uri_base, "gom5_201812.nc")
x <- nc_open(uri)
x
```

# Data structure

## Nodes and Elements

The mesh is defined by non-intersecting triangular elements bounded by
three nodes. Nodes are shown below as the solid dots, three nodes define
the boundary of element. Scalar values, like temperature, salinty and
height are defined at nodes. Vector values, such as velocity are defined
at the element centroids. Within the NetCDF object you can make complete
identification by examining which variables use the `node` dimensions
versus those that use the `nele` dimension.

![*Fig. 3.11 **from the user manual**: Schematic of the control volume
used to calculate scalar variables and vertical velocity used in FVCOM.
F is a general symbol representing scalar variables such as zeta, T, S,
Km, Kh, and vertical velocity. A **solid dot** is the node of the
triangles where scalar variable or vertical velocity is calculated and a
**crossed open circle** is the centroid of a triangle where the
horizontal velocity is calculated.*](inst/images/nodes-elements.png)

## Node and Element locations

Use the functions `fvcom_nodes` and `fvcom_elems` to extract location
information as either `xy` or `lonlat` pairs. Note we get both xy and
lonlat here for nodes. The same can be had for elements.

```{r}
dplyr::left_join(fvcom::fvcom_nodes(x, what = 'lonlat'),                   
                 fvcom::fvcom_nodes(x, what = 'xy'), by = "node") 
```

### Timestamps

Retrieving the timestamps provides you with a choice for assumptions you
want to make about the hourly model output. Timestamps stored internally
do not land on each hour.

```{r}
fvcom_time(x, internal = TRUE) |> head()
```

We can get a proxy for these times, but *prettily* settled on each hour.
The choice is yours.

```{r}
fvcom_time(x, internal = FALSE) |> head()
```

##  Variables

Variables (the oceanographic ones) can be extract by node or element. It
is possible to select a subset, but the operation on the whole dataset
is quick and it is just as easy to subset after you have the table in
hand.

```{r}
v <- get_node_var(x, var = 'zeta')
v
```

## Mesh

Some computation is required (not a lot) to produce the mesh which is
comprised of non-intersecting polygons (triangles). Meshes can be
derived from the lists of nodes or the list of elements. We have chosen
to use elements by default as they are simple to construct from the
adjancency lists provided in the NetCDF resource.

The each element in the mesh is defined by the three nodes identified by
index, “p1”, “p2” and “p3”. The geometry is defined by either “lonlat”
coordinates or projected mercator coordinates, “xy”.

```{r}
mesh <- get_mesh_geometry(x, what = 'lonlat') 
mesh
```

```{r}
plot(sf::st_geometry(mesh), axes = TRUE)
```


### Mesh with variables

We can assign variable values to the polygons by reusing the mesh table.
These variables are associated with either node or elements. If
node-referenced variables are requested the mean of three neighboring
nodes (which define an element) is computed. Where element-referenced
variables are requested no averaging is done - the values are simply
assigned to the mesh element.

```{r}
mesh <- get_mesh(x, vars = c("zeta", "u", "v"), mesh = mesh)
plot(mesh[c("u", "v")], lty = 'blank', main = c("u", "v"))
```


You can request variables at various dimensions such as times and sigma
levels/layers - the default is the first of each dimension. While you
can request one of these in ‘real’ values (such as `POSIXct` time), you
can also provide a 1-based index into that dimension. The example below
requests the same variables as above but at the 24th time interval. See
the functions `get_node_var` and `get_elem_var` for details.

```{r}
mesh <- get_mesh(x, vars = c("zeta", "u", "v"), mesh = mesh, time = 24)
plot(mesh[c("u", "v")], lty = 'blank', main = c("u", "v"))
```

### Rasterize

The mesh can be interpolated on to a regular grid (“rasterize”).

```{r}
template = default_template(mesh)
uv <- sapply(c("u", "v"), 
    function(f) {
        fvcom::rasterize(mesh[f], template = template) 
    }, simplify = FALSE)
par(mfrow = c(1,2))
plot(uv[['u']], key.pos = NULL, reset = FALSE)
plot(uv[['v']], key.pos = NULL, reset = FALSE)
```



```{r}
ncdf4::nc_close(x)
```