---
title: "FVCOM"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Convenient access to [FVCOM](http://fvcom.smast.umassd.edu/fvcom/)
datasets from R.

### Background

[FVCOM](http://fvcom.smast.umassd.edu/fvcom/) produces ocean circulation
models on an irregular mesh. Here is the [user
manual](http://fvcom.smast.umassd.edu/wp-content/uploads/2013/11/MITSG_12-25.pdf).

As an example, [NECOFS](http://fvcom.smast.umassd.edu/necofs/) leverages
this mesh to deliver models for the Gulf of Maine, Mass coastal waters
and a small number of estuaries. NECOFS provides an [interactive web
map](http://134.88.228.119:8080/fvcomwms/)

## Requirements

 + [R 4.1+]()
  + [grDevices]()
  + [graphics]()
  + [stats]()
  + [rlang]()
  + [ncdf4]()
  + [tibble]()
  + [dplyr]()
  + [raster]()
  + [sf]()
  + [fasterize]()
  + [locate]() # Not available on CRAN

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
